---
output: word_document
---
###Simulated sampling Dynamic modeling

```{r}
library("unmarked")
```

#First we simulate our detection history for 30 sites with four primary sampling periods, and three secondary sampling periods each.
```{r, echo=FALSE}
s1.1a<-rbinom(n=10, size=1, prob=0.7)
s1.1b<-rbinom(n=10, size=1, prob=0.4)
s1.1c<-rbinom(n=10, size=1, prob=0.2)
s1.2a<-rbinom(n=10, size=1, prob=0.7)
s1.2b<-rbinom(n=10, size=1, prob=0.4)
s1.2c<-rbinom(n=10, size=1, prob=0.2)
s1.3a<-rbinom(n=10, size=1, prob=0.7)
s1.3b<-rbinom(n=10, size=1, prob=0.4)
s1.3c<-rbinom(n=10, size=1, prob=0.2)

s1.1 <-c(s1.1a,s1.1b,s1.1c)
s1.2 <-c(s1.2a,s1.2b,s1.2c)
s1.3 <-c(s1.3a,s1.3b,s1.3c)

s2.1a<-rbinom(n=10, size=1, prob=0.7)
s2.1b<-rbinom(n=10, size=1, prob=0.4)
s2.1c<-rbinom(n=10, size=1, prob=0.2)
s2.2a<-rbinom(n=10, size=1, prob=0.7)
s2.2b<-rbinom(n=10, size=1, prob=0.4)
s2.2c<-rbinom(n=10, size=1, prob=0.2)
s2.3a<-rbinom(n=10, size=1, prob=0.7)
s2.3b<-rbinom(n=10, size=1, prob=0.4)
s2.3c<-rbinom(n=10, size=1, prob=0.2)

s2.1 <-c(s2.1a,s2.1b,s2.1c)
s2.2 <-c(s2.2a,s2.2b,s2.2c)
s2.3 <-c(s2.3a,s2.3b,s2.3c)

s3.1a<-rbinom(n=10, size=1, prob=0.9)
s3.1b<-rbinom(n=10, size=1, prob=0.4)
s3.1c<-rbinom(n=10, size=1, prob=0.1)
s3.2a<-rbinom(n=10, size=1, prob=0.9)
s3.2b<-rbinom(n=10, size=1, prob=0.4)
s3.2c<-rbinom(n=10, size=1, prob=0.2)
s3.3a<-rbinom(n=10, size=1, prob=0.9)
s3.3b<-rbinom(n=10, size=1, prob=0.4)
s3.3c<-rbinom(n=10, size=1, prob=0.1)

s3.1 <-c(s3.1a,s3.1b,s3.1c)
s3.2 <-c(s3.2a,s3.2b,s3.2c)
s3.3 <-c(s3.3a,s3.3b,s3.3c)

s4.1a<-rbinom(n=10, size=1, prob=0.9)
s4.1b<-rbinom(n=10, size=1, prob=0.4)
s4.1c<-rbinom(n=10, size=1, prob=0.1)
s4.2a<-rbinom(n=10, size=1, prob=0.9)
s4.2b<-rbinom(n=10, size=1, prob=0.4)
s4.2c<-rbinom(n=10, size=1, prob=0.1)
s4.3a<-rbinom(n=10, size=1, prob=0.9)
s4.3b<-rbinom(n=10, size=1, prob=0.4)
s4.3c<-rbinom(n=10, size=1, prob=0.1)

s4.1 <-c(s4.1a,s4.1b,s4.1c)
s4.2 <-c(s4.2a,s4.2b,s4.2c)
s4.3 <-c(s4.3a,s4.3b,s4.3c)
sampling.ocup<- cbind(s1.1,s1.2,s1.3,s2.1,s2.2,s2.3,s3.1,s3.2,s3.3,s4.1,s4.2,s4.3)
```

```{r}
View(sampling.ocup)
```
This simulated data has some underlying charachteristics:

**environment a (top 10 rows, the best environment for bats, also occupancy increases with time)**
```{r}
mean(sampling.ocup[1:10])
```
**environment b (rows 11 to 20, medium environment, occupancy stays the same)**
```{r}
mean(sampling.ocup[11:20])
```
**environment c (rows 21 to 30) poor environment for bats, also there is extintion)**
```{r}
mean(sampling.ocup[21:30])
```

Now we will simulate some variables for the site covariates.
#Site cov static
**the more variable 1, the better for bats**

```{r, echo=FALSE}
v.1a<-rnorm(n=10, mean=20, sd=0.7)
v.1b<-rnorm(n=10, mean=10, sd=0.7)
v.1c<-rnorm(n=10, mean=5, sd=0.7)

v.1 <-c(v.1a,v.1b,v.1c)
```

```{r}
mean(v.1[1:10])
mean(v.1[11:20])
mean(v.1[21:30])
```


#the less variable 2 better for bats
```{r,echo=FALSE}
v.2a<-rnorm(n=10, mean=20, sd=0.7)
v.2b<-rnorm(n=10, mean=40, sd=0.7)
v.2c<-rnorm(n=10, mean=60, sd=0.7)

v.2 <-c(v.2a,v.2b,v.2c)
```

```{r}
mean(v.2[1:10])
mean(v.2[11:20])
mean(v.2[21:30])
```

#variable 3 does not mater to bats

```{r,echo=FALSE}
v.3a<-rnorm(n=10, mean=20, sd=0.7)
v.3b<-rnorm(n=10, mean=20, sd=0.7)
v.3c<-rnorm(n=10, mean=20, sd=0.7)


v.3 <-c(v.3a,v.3b,v.3c)
```

```{r}
mean(v.3[1:10])
mean(v.3[11:20])
mean(v.3[21:30])
```

```{r}
sampling.cov<- cbind(v.1,v.2, v.3)
```
#yearly colonization extintion variable

```{r}
temp1<-rnorm(n=30, mean=20, sd=10)
temp2<-rnorm(n=30, mean=40, sd=10)
temp3<-rnorm(n=30, mean=50, sd=10)
temp4<-rnorm(n=30, mean=60, sd=10)


temps <-cbind(temp1,temp2, temp3, temp4)
```
#observer within secondary it could be variable


#observers dont vary in the model, they should all be the same

```{r}
obs1<-rnorm(n=30, mean=40, sd=0.7)
obs2<-rnorm(n=30, mean=40, sd=0.7)
obs3<-rnorm(n=30, mean=40, sd=0.7)
obs4<-rnorm(n=30, mean=40, sd=0.7)
obs5<-rnorm(n=30, mean=40, sd=0.7)
obs6<-rnorm(n=30, mean=40, sd=0.7)
obs7<-rnorm(n=30, mean=40, sd=0.7)
obs8<-rnorm(n=30, mean=40, sd=0.7)
obs9<-rnorm(n=30, mean=40, sd=0.7)
obs10<-rnorm(n=30, mean=40, sd=0.7)
obs11<-rnorm(n=30, mean=40, sd=0.7)
obs12<-rnorm(n=30, mean=40, sd=0.7)

observers1<-data.frame(cbind(obs1,obs2,obs3, obs4, obs5, obs6, obs7, obs8, obs9, obs10, obs11, obs12))

observers2<-data.frame(cbind(obs1,obs2,obs3, obs4, obs5, obs6, obs7, obs8, obs9, obs10, obs11, obs12))

observers<-list(observers1, observers2)
names(observers) <-c("obs1", "obs2")
```
#primary model
```{r}
umf <- unmarkedMultFrame(y = sampling.ocup, 
                            siteCovs = data.frame(sampling.cov), 
                            yearlySiteCovs=data.frame(temps),
                            obsCovs=observers, numPrimary=3)
```
#Dynamic model
#first term static variables
#second term colonization (variable)
#third extintion (variable)
#detection (observer)
#the best model should take into acount v.1 and v.2, but not v.3, it shouldn't take into acount observers

```{r}
model1 <- colext(~v.1+v.2+v.3, ~temp1, ~temp1, ~obs1+obs2, umf)

model2 <- colext(~1, ~1, ~1, ~1, umf)

model3 <- colext(~v.1+v.2, ~temp1, ~temp1, ~1, umf) #this should be the best model
model4 <-colext(~v.1+v.2, ~1, ~1, ~1, umf)
```
#graph
