
#Sampling Design
###author: Derek Corcoran
####Last update: `r Sys.Date() `
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

First we load the spatial packages

```{r, message=FALSE, warning=FALSE}
library("raster", lib.loc="~/R/win-library/3.2")
library("rasterVis", lib.loc="~/R/win-library/3.2")
library("maps", lib.loc="~/R/win-library/3.2")
library("maptools", lib.loc="~/R/win-library/3.2")
library("rgdal", lib.loc="~/R/win-library/3.2")
```
Then we read the needed rasters

```{r, message=FALSE, warning=FALSE}
PNF<- readGDAL("C:/Users/usuario/Bats_California/layers/PNF.asc")
PNF<-raster (PNF)
plot(PNF)
bc <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_canopy.asc")
bc<-raster (bc)
plot(bc)
bb <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_basal.asc")
bb<-raster (bb)
plot(bb)
bs <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_severity.asc")
bs<-raster (bs)
plot(bs)
topo <- readGDAL("C:/Users/usuario/Bats_California/layers/plumastopo.asc")
topo<-raster (topo)
plot(topo)
Vegetation_existing <- readGDAL("C:/Users/usuario/Bats_California/layers/Vegetation_existing.asc")
Vegetation_existing<-raster (Vegetation_existing)
plot(Vegetation_existing)
FireReturnIntervalDeparture <- readGDAL("C:/Users/usuario/Bats_California/layers/FireReturnIntervalDeparture.asc")
FireReturnIntervalDeparture<-raster (FireReturnIntervalDeparture)
plot(FireReturnIntervalDeparture)
TreatmentsStorrie <- readGDAL("C:/Users/usuario/Bats_California/layers/TreatmentsStorrie.asc")
TreatmentsStorrie<-raster (TreatmentsStorrie)
plot(TreatmentsStorrie)
```

#Change outlayers and extract NAs 

In order to classify the raster we will get rid of unnecesary outlayers, and change NAs to 0

```{r}
df.bb <- data.frame(id=c(NA,1,2,3,4,5,6,7,255), v=c(0,1,2,3,4,5,6,7,8))
bb1 <- subs(bb, df.bb,subswithNA=FALSE)
df.bs <- data.frame(id=c(NA,1,2,3,4,255), v=c(0,1,2,3,4,5))
bs1 <- subs(bs, df.bs,subswithNA=FALSE)
df.bc <- data.frame(id=c(NA,1,2,3,4,5,255), v=c(0,1,2,3,4,5,6))
bc1 <- subs(bc, df.bc,subswithNA=FALSE)
```

#Put all rasters in the same projection

```{r}
bb<-projectRaster(bb1, PNF)
bs<-projectRaster(bs1, PNF)
bc<-projectRaster(bc1, PNF)
Vegetation_existing<-projectRaster(Vegetation_existing, PNF)
FireReturnIntervalDeparture<-projectRaster(FireReturnIntervalDeparture, PNF)
TreatmentsStorrie<-projectRaster(TreatmentsStorrie, PNF)

```
#Put them all in the same resolution and size

```{r}
bc<-resample(bc, PNF)
bb<-resample(bb, PNF)
bs<-resample(bs, PNF)
Vegetation_existing<-resample(Vegetation_existing, PNF)
FireReturnIntervalDeparture<-resample(FireReturnIntervalDeparture, PNF)
TreatmentsStorrie<-resample(TreatmentsStorrie, PNF)
topo<-resample(topo,PNF)
```

#Prepare a distance from river/road raster

```{r}
roads.v <- readOGR(dsn="C:/Users/usuario/Bats_California/layers",layer="Roads")
roads.v <- spTransform(roads.v, CRS("+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"))
plot(PNF)
lines(roads.v)
template <- PNF  # this will be the template
template[] <- NA  # assigns all values as NA
roads.r <- rasterize(roads.v, template, field=1)
summary(roads.r)          # pixels crossed by a road have "1" 
plot(roads.r, add=TRUE)
```
after rasterizing the roads we make the new raster
```{r}
roaddist.r <- distance(roads.r)
class(roaddist.r)
# Check:
plot(roaddist.r)
lines(roads.v)
```
#Check for correlation between rasters

```{r}
AllLayers <-stack(topo,TreatmentsStorrie,FireReturnIntervalDeparture, Vegetation_existing, roaddist.r)
plot (AllLayers, colNA="black")
pairs(AllLayers)
AllLayers[is.na(AllLayers)] <- 0
plot(AllLayers)
```

#Clasification example

Even though we shouldn't classify using 3 layers of such high classification we will use the RasterBrick of the three burn classifications to exemplify how we will divide the area into areas of similar characteristics.
Here we will ask R to use kmeans to sort the area into 3 types of habitat using the abovementioned rasterbrick:

#now with every layer

```{r, echo=FALSE}
valuetable2 <- getValues(AllLayers)
head(valuetable2) 
summary(valuetable2)
km <- kmeans(na.omit(valuetable2), centers = 5, iter.max = 100, nstart = 10)
# create a blank raster with default values of 0
rNA <- setValues(raster(AllLayers), 0)
for(i in 1:nlayers(AllLayers)){
  rNA[is.na(AllLayers[[i]])] <- 1
}
# convert rNA to an integer vector
rNA <- getValues(rNA)
# convert valuetable to a data.frame
valuetable2 <- as.data.frame(valuetable2)
# if rNA is a 0, assign the cluster value at that position
valuetable2$class[rNA==0] <- km$cluster
# if rNA is a 1, assign an NA at that position
valuetable2$class[rNA==1] <- NA
# create a blank raster
classes <- raster(AllLayers)
# assign values from the 'class' column of valuetable
classes2 <- setValues(classes, valuetable2$class)
plot(classes2, legend=TRUE)
summary (classes2)
```

More info on how to do this clasification in *https://geoscripting-wur.github.io/AdvancedRasterAnalysis/*

#Extract Random points from each habitat type

```{r}
set.seed(10)
df.class.4 <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,NA,4,NA))
class4 <- subs(classes2, df.class.4,subswithNA=FALSE)
points4<-sampleRandom(class4,10, na.rm=TRUE, xy=TRUE)
df.class.3 <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,3,NA,NA))
class3 <- subs(classes2, df.class.3,subswithNA=FALSE)
points3<-sampleRandom(class3,10, na.rm=TRUE, xy=TRUE)
plot(classes2)
points (points4, col= "black")
points (points3, col= "red")
```