---
output: word_document
---

#Sampling Design
###author: Derek Corcoran
####Last update: `r Sys.Date() `
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

First we load the spatial packages

```{r}
library("raster", lib.loc="~/R/win-library/3.2")
library("rasterVis", lib.loc="~/R/win-library/3.2")
library("maps", lib.loc="~/R/win-library/3.2")
library("maptools", lib.loc="~/R/win-library/3.2")
library("rgdal", lib.loc="~/R/win-library/3.2")
```
read rasters

```{r}
bc <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_canopy.asc")
bc<-raster (bc)
plot(bc)
bb <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_basal.asc")
bb<-raster (bb)
plot(bb)
bs <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_severity.asc")
bs<-raster (bs)
plot(bs)
topo <- readGDAL("C:/Users/usuario/Bats_California/layers/plumastopo.asc")
topo<-raster (topo)
plot(topo)
summary(bs)
```

#Change outlayers and extract NAs 

In order to classify the raster we will get rid of unnecesary outlayers, and change NAs to 0

```{r}
df.bb <- data.frame(id=c(NA,1,2,3,4,5,6,7,255), v=c(0,1,2,3,4,5,6,7,8))
bb <- subs(bb, df.bb,subswithNA=FALSE)
df.bs <- data.frame(id=c(NA,1,2,3,4,255), v=c(0,1,2,3,4,5))
bs <- subs(bs, df.bs,subswithNA=FALSE)
df.bc <- data.frame(id=c(NA,1,2,3,4,5,255), v=c(0,1,2,3,4,5,6))
bc <- subs(bc, df.bc,subswithNA=FALSE)
```

#Check for correlation between rasters

```{r}
burn <-brick(bc, bs, bb)
plot (burn)
layerStats(burn, "pearson",na.rm=TRUE)
```

#Clasification example

Even though we shouldn't classify using 3 layers of such high classification we will use the RasterBrick of the three burn classifications to exemplify how we will divide the area into areas of similar characteristics.
Here we will ask R to use kmeans to sort the area into 3 types of habitat using the abovementioned rasterbrick:

```{r, echo=FALSE}
valuetable <- getValues(burn)
head(valuetable)
km <- kmeans(na.omit(valuetable), centers = 3, iter.max = 100, nstart = 10)
# create a blank raster with default values of 0
rNA <- setValues(raster(burn), 0)
for(i in 1:nlayers(burn)){
  rNA[is.na(burn[[i]])] <- 1
}
# convert rNA to an integer vector
rNA <- getValues(rNA)
# convert valuetable to a data.frame
valuetable <- as.data.frame(valuetable)
# if rNA is a 0, assign the cluster value at that position
valuetable$class[rNA==0] <- km$cluster
# if rNA is a 1, assign an NA at that position
valuetable$class[rNA==1] <- NA
# create a blank raster
classes <- raster(burn)
# assign values from the 'class' column of valuetable
classes <- setValues(classes, valuetable$class)
plot(classes, legend=FALSE,col=c("dark green", "orange", "light blue"))
```

More info on how to do this clasification in *https://geoscripting-wur.github.io/AdvancedRasterAnalysis/*