
#Sampling Design
###author: Derek Corcoran
####Last update: `r Sys.Date() `
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

First we load the spatial packages

```{r, message=FALSE, warning=FALSE}
library("raster", lib.loc="~/R/win-library/3.2")
library("rasterVis", lib.loc="~/R/win-library/3.2")
library("maps", lib.loc="~/R/win-library/3.2")
library("maptools", lib.loc="~/R/win-library/3.2")
library("rgdal", lib.loc="~/R/win-library/3.2")
```
Then we read the needed rasters

```{r, message=FALSE, warning=FALSE}
PNF<- readGDAL("C:/Users/usuario/Bats_California/layers/PNF.asc")
PNF<-raster (PNF)
plot(PNF)
bc <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_canopy.asc")
bc<-raster (bc)
plot(bc)
bb <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_basal.asc")
bb<-raster (bb)
plot(bb)
bs <- readGDAL("C:/Users/usuario/Bats_California/layers/burn_severity.asc")
bs<-raster (bs)
plot(bs)
topo <- readGDAL("C:/Users/usuario/Bats_California/layers/plumastopo.asc")
topo<-raster (topo)
plot(topo)
Vegetation_existing <- readGDAL("C:/Users/usuario/Bats_California/layers/Vegetation_existing.asc")
Vegetation_existing<-raster (Vegetation_existing)
plot(Vegetation_existing)
FireReturnIntervalDeparture <- readGDAL("C:/Users/usuario/Bats_California/layers/FireReturnIntervalDeparture.asc")
FireReturnIntervalDeparture<-raster (FireReturnIntervalDeparture)
plot(FireReturnIntervalDeparture)
TreatmentsStorrie <- readGDAL("C:/Users/usuario/Bats_California/layers/TreatmentsStorrie.asc")
TreatmentsStorrie<-raster (TreatmentsStorrie)
plot(TreatmentsStorrie)
```

#Change outlayers and extract NAs 

In order to classify the raster we will get rid of unnecesary outlayers, and change NAs to 0

```{r}
df.bb <- data.frame(id=c(NA,1,2,3,4,5,6,7,255), v=c(0,1,2,3,4,5,6,7,8))
bb1 <- subs(bb, df.bb,subswithNA=FALSE)
df.bs <- data.frame(id=c(NA,1,2,3,4,255), v=c(0,1,2,3,4,5))
bs1 <- subs(bs, df.bs,subswithNA=FALSE)
df.bc <- data.frame(id=c(NA,1,2,3,4,5,255), v=c(0,1,2,3,4,5,6))
bc1 <- subs(bc, df.bc,subswithNA=FALSE)
lidar.limit <- readOGR(dsn="C:/Users/usuario/Bats_California/layers",layer="LiDAR_Boundary")
lidar.limit <- spTransform(lidar.limit, CRS("+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"))
plot(PNF)
plot(lidar.limit, add=TRUE)
```

#Put all rasters in the same projection

```{r}
bb<-projectRaster(bb1, PNF)
bs<-projectRaster(bs1, PNF)
bc<-projectRaster(bc1, PNF)
Vegetation_existing<-projectRaster(Vegetation_existing, PNF)
FireReturnIntervalDeparture<-projectRaster(FireReturnIntervalDeparture, PNF)
TreatmentsStorrie<-projectRaster(TreatmentsStorrie, PNF)

```
#Put them all in the same resolution and size

```{r}
bc<-resample(bc, PNF)
bb<-resample(bb, PNF)
bs<-resample(bs, PNF)
Vegetation_existing<-resample(Vegetation_existing, PNF)
FireReturnIntervalDeparture<-resample(FireReturnIntervalDeparture, PNF)
FireReturnIntervalDeparture[is.na(FireReturnIntervalDeparture)] <- 0
TreatmentsStorrie<-resample(TreatmentsStorrie, PNF)
topo<-resample(topo,PNF)
TreatmentsStorrie[is.na(TreatmentsStorrie)] <- 0

```

#Prepare a distance from river/road raster

```{r}
roads.v <- readOGR(dsn="C:/Users/usuario/Bats_California/layers",layer="Roads")
roads.v <- spTransform(roads.v, CRS("+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"))
plot(PNF)
lines(roads.v)
template <- PNF  # this will be the template
template[] <- NA  # assigns all values as NA
roads.r <- rasterize(roads.v, template, field=1)
summary(roads.r)          # pixels crossed by a road have "1" 
plot(roads.r, add=TRUE)
```
after rasterizing the roads we make the new raster
```{r}
roaddist.r <- distance(roads.r)
roaddist.r<- roaddist.r*PNF
class(roaddist.r)
# Check:
plot(roaddist.r)
lines(roads.v)
```

```{r}
plot(roaddist.r)
summary(roaddist.r)
```
#Check for correlation between rasters

```{r}
AllLayers1 <-stack(topo,TreatmentsStorrie,FireReturnIntervalDeparture, Vegetation_existing,roaddist.r)
AllLayers2<-AllLayers1*PNF
plot (AllLayers2, colNA="black")
pairs(AllLayers2)
rasterRescale<-function(r){
    ((r-cellStats(r,"min"))/(cellStats(r,"max")-cellStats(r,"min")))
}
AllLayers3<-rasterRescale(AllLayers2)
plot(AllLayers3)
```

#Clasification example

Even though we shouldn't classify using 3 layers of such high classification we will use the RasterBrick of the three burn classifications to exemplify how we will divide the area into areas of similar characteristics.
Here we will ask R to use kmeans to sort the area into 3 types of habitat using the abovementioned rasterbrick:


#now with every layer


```{r, echo=FALSE}
valuetable <- getValues(AllLayers3)
head(valuetable) 
summary(valuetable)
km1 <- kmeans(na.omit(valuetable), centers = 5, iter.max = 100, nstart = 10)
# create a blank raster with default values of 0
rNA <- setValues(raster(AllLayers3), 0)
for(i in 1:nlayers(AllLayers3)){
  rNA[is.na(AllLayers3[[i]])] <- 1
}
# convert rNA to an integer vector
rNA <- getValues(rNA)
# convert valuetable to a data.frame
valuetable <- as.data.frame(valuetable)
# if rNA is a 0, assign the cluster value at that position
valuetable$class[rNA==0] <- km1$cluster
# if rNA is a 1, assign an NA at that position
valuetable$class[rNA==1] <- NA
# create a blank raster
classes2 <- raster(AllLayers3)
# assign values from the 'class' column of valuetable
classes2a <- setValues(classes2, valuetable2$class)
plot(classes2a, legend=TRUE, colNA="black")
summary (classes2a)
```

More info on how to do this clasification in *https://geoscripting-wur.github.io/AdvancedRasterAnalysis/*

#separate layers acording to places with or without fire

```{r}
df.bs1 <- data.frame(id=c(0,1,2,3,4,5), v=c(NA,1,1,1,1,1))
fire <- subs(bs1, df.bs1,subswithNA=FALSE)
fire<-projectRaster(fire, PNF)

df.bs2 <- data.frame(id=c(NA,1), v=c(1,NA))
not.fire <- subs(fire, df.bs2,subswithNA=FALSE)

plot(fire, colNA="black")
plot(not.fire, colNA="black")
plot(lidar.limit,add=TRUE)
summary(fire)
summary(not.fire)
classes.with.fire<-classes2a*fire
classes.without.fire<-classes2a*not.fire

plot(classes.with.fire,colNA="black")
summary(classes.with.fire)
plot(classes.without.fire,colNA="black")
summary(classes.without.fire)
```

#Extract Random points from each habitat type with equal number in fire and non fire


```{r}
set.seed(11)

df.class.1f <- data.frame(id=c(1,2,3,4,5), v=c(1,NA,NA,NA,NA))
class1f <- subs(classes.with.fire, df.class.1f,subswithNA=FALSE)
points1f<-sampleRandom(class1f,3, na.rm=TRUE, xy=TRUE)

df.class.1nf <- data.frame(id=c(1,2,3,4,5), v=c(1,NA,NA,NA,NA))
class1nf <- subs(classes.without.fire, df.class.1nf,subswithNA=FALSE)
points1nf<-sampleRandom(class1nf,3, na.rm=TRUE, xy=TRUE)

df.class.2f <- data.frame(id=c(1,2,3,4,5), v=c(NA,2,NA,NA,NA))
class2f <- subs(classes.with.fire, df.class.2f,subswithNA=FALSE)
points2f<-sampleRandom(class2f,3, na.rm=TRUE, xy=TRUE)

df.class.2nf <- data.frame(id=c(1,2,3,4,5), v=c(NA,2,NA,NA,NA))
class2nf <- subs(classes.without.fire, df.class.2nf,subswithNA=FALSE)
points2nf<-sampleRandom(class2nf,3, na.rm=TRUE, xy=TRUE)

df.class.3f <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,3,NA,NA))
class3f <- subs(classes.with.fire, df.class.3f,subswithNA=FALSE)
points3f<-sampleRandom(class3f,3, na.rm=TRUE, xy=TRUE)

df.class.3nf <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,3,NA,NA))
class3nf <- subs(classes.without.fire, df.class.3nf,subswithNA=FALSE)
points3nf<-sampleRandom(class3nf,3, na.rm=TRUE, xy=TRUE)

df.class.4f <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,NA,4,NA))
class4f <- subs(classes.with.fire, df.class.4f,subswithNA=FALSE)
points4f<-sampleRandom(class4f,3, na.rm=TRUE, xy=TRUE)

df.class.4nf <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,NA,4,NA))
class4nf <- subs(classes.without.fire, df.class.4nf,subswithNA=FALSE)
points4nf<-sampleRandom(class4nf,3, na.rm=TRUE, xy=TRUE)

df.class.5f <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,NA,NA,5))
class5f <- subs(classes.with.fire, df.class.5f,subswithNA=FALSE)
points5f<-sampleRandom(class5f,3, na.rm=TRUE, xy=TRUE)

df.class.5nf <- data.frame(id=c(1,2,3,4,5), v=c(NA,NA,NA,NA,5))
class5nf <- subs(classes.without.fire, df.class.5nf,subswithNA=FALSE)
points5nf<-sampleRandom(class5nf,3, na.rm=TRUE, xy=TRUE)

plot(classes2a, colNA="white")
points (points1f, col= "black", pch=21, bg="black")
points (points1nf, col="black",pch=24,bg="black")
points (points2f, col= "blue",pch=21,bg="blue")
points (points2nf, col="blue", pch=24, bg="blue")
points (points3f, col= "red",pch=21, bg="red")
points (points3nf, col="red", pch=24,bg="red")
points (points4f, col= "brown",pch=21,bg="brown")
points (points4nf, col="brown", pch=24,bg="brown")
points (points5f, col= "purple",pch=21,bg="purple")
points (points5nf, col="purple", pch=24, bg="purple")
```


#Simulated sampling Dynamic modeling

```{r}
library("unmarked", lib.loc="~/R/win-library/3.2")
```

###First we simulate our detection history for 30 sites with four primary sampling periods, and three secondary sampling periods each.
```{r, echo=FALSE}
set.seed(20)
s1.1a<-rbinom(n=10, size=1, prob=0.9)
s1.1b<-rbinom(n=10, size=1, prob=0.4)
s1.1c<-rbinom(n=10, size=1, prob=0.2)
s1.2a<-rbinom(n=10, size=1, prob=0.9)
s1.2b<-rbinom(n=10, size=1, prob=0.4)
s1.2c<-rbinom(n=10, size=1, prob=0.2)
s1.3a<-rbinom(n=10, size=1, prob=0.9)
s1.3b<-rbinom(n=10, size=1, prob=0.4)
s1.3c<-rbinom(n=10, size=1, prob=0.2)

s1.1 <-c(s1.1a,s1.1b,s1.1c)
s1.2 <-c(s1.2a,s1.2b,s1.2c)
s1.3 <-c(s1.3a,s1.3b,s1.3c)

s2.1a<-rbinom(n=10, size=1, prob=0.9)
s2.1b<-rbinom(n=10, size=1, prob=0.4)
s2.1c<-rbinom(n=10, size=1, prob=0.2)
s2.2a<-rbinom(n=10, size=1, prob=0.9)
s2.2b<-rbinom(n=10, size=1, prob=0.4)
s2.2c<-rbinom(n=10, size=1, prob=0.2)
s2.3a<-rbinom(n=10, size=1, prob=0.9)
s2.3b<-rbinom(n=10, size=1, prob=0.4)
s2.3c<-rbinom(n=10, size=1, prob=0.2)

s2.1 <-c(s2.1a,s2.1b,s2.1c)
s2.2 <-c(s2.2a,s2.2b,s2.2c)
s2.3 <-c(s2.3a,s2.3b,s2.3c)

s3.1a<-rbinom(n=10, size=1, prob=0.9)
s3.1b<-rbinom(n=10, size=1, prob=0.4)
s3.1c<-rbinom(n=10, size=1, prob=0.1)
s3.2a<-rbinom(n=10, size=1, prob=0.9)
s3.2b<-rbinom(n=10, size=1, prob=0.4)
s3.2c<-rbinom(n=10, size=1, prob=0.2)
s3.3a<-rbinom(n=10, size=1, prob=0.9)
s3.3b<-rbinom(n=10, size=1, prob=0.4)
s3.3c<-rbinom(n=10, size=1, prob=0.1)

s3.1 <-c(s3.1a,s3.1b,s3.1c)
s3.2 <-c(s3.2a,s3.2b,s3.2c)
s3.3 <-c(s3.3a,s3.3b,s3.3c)

s4.1a<-rbinom(n=10, size=1, prob=0.9)
s4.1b<-rbinom(n=10, size=1, prob=0.4)
s4.1c<-rbinom(n=10, size=1, prob=0.1)
s4.2a<-rbinom(n=10, size=1, prob=0.9)
s4.2b<-rbinom(n=10, size=1, prob=0.4)
s4.2c<-rbinom(n=10, size=1, prob=0.1)
s4.3a<-rbinom(n=10, size=1, prob=0.9)
s4.3b<-rbinom(n=10, size=1, prob=0.4)
s4.3c<-rbinom(n=10, size=1, prob=0.1)

s4.1 <-c(s4.1a,s4.1b,s4.1c)
s4.2 <-c(s4.2a,s4.2b,s4.2c)
s4.3 <-c(s4.3a,s4.3b,s4.3c)
sampling.ocup1<- cbind(s1.1,s1.2,s1.3,s2.1,s2.2,s2.3,s3.1,s3.2,s3.3,s4.1,s4.2,s4.3)
```
```{r,echo=FALSE}
library("knitr", lib.loc="~/R/win-library/3.2")
kable(sampling.ocup1)
```

This simulated data has some underlying characteristics:

**environment a (top 10 rows, the best environment for bats, also occupancy increases with time)**
```{r}
mean(sampling.ocup1[1:10])
```
**environment b (rows 11 to 20, medium environment, occupancy stays the same)**
```{r}
mean(sampling.ocup1[11:20])
```
**environment c (rows 21 to 30) poor environment for bats, also there is extintion)**
```{r}
mean(sampling.ocup1[21:30])
```

Now we will simulate some variables for the site covariates.
###Site cov static
**the more variable 1, the better for bats**

```{r, echo=FALSE}
v.1a<-rnorm(n=10, mean=20, sd=0.7)
v.1b<-rnorm(n=10, mean=10, sd=0.7)
v.1c<-rnorm(n=10, mean=5, sd=0.7)

v.1 <-c(v.1a,v.1b,v.1c)
```

```{r}
mean(v.1[1:10])
mean(v.1[11:20])
mean(v.1[21:30])
```


###the less variable 2 better for bats
```{r,echo=FALSE}
v.2a<-rnorm(n=10, mean=20, sd=0.7)
v.2b<-rnorm(n=10, mean=40, sd=0.7)
v.2c<-rnorm(n=10, mean=60, sd=0.7)

v.2 <-c(v.2a,v.2b,v.2c)
```

```{r}
mean(v.2[1:10])
mean(v.2[11:20])
mean(v.2[21:30])
```

###variable 3 does not mater to bats

```{r,echo=FALSE}
v.3a<-rnorm(n=10, mean=20, sd=0.7)
v.3b<-rnorm(n=10, mean=20, sd=0.7)
v.3c<-rnorm(n=10, mean=20, sd=0.7)


v.3 <-c(v.3a,v.3b,v.3c)
```

```{r}
mean(v.3[1:10])
mean(v.3[11:20])
mean(v.3[21:30])
```

```{r}
sampling.cov<- cbind(v.1,v.2, v.3)
```
###yearly colonization extintion variable

```{r}
temp1<-rnorm(n=30, mean=20, sd=10)
temp2<-rnorm(n=30, mean=40, sd=10)
temp3<-rnorm(n=30, mean=50, sd=10)
temp4<-rnorm(n=30, mean=60, sd=10)


temps <-cbind(temp1,temp2, temp3, temp4)
```
###observer within secondary it could be variable


###observers dont vary in the model, they should all be the same

```{r}
obs1<-rnorm(n=30, mean=40, sd=0.7)
obs2<-rnorm(n=30, mean=40, sd=0.7)
obs3<-rnorm(n=30, mean=40, sd=0.7)
obs4<-rnorm(n=30, mean=40, sd=0.7)
obs5<-rnorm(n=30, mean=40, sd=0.7)
obs6<-rnorm(n=30, mean=40, sd=0.7)
obs7<-rnorm(n=30, mean=40, sd=0.7)
obs8<-rnorm(n=30, mean=40, sd=0.7)
obs9<-rnorm(n=30, mean=40, sd=0.7)
obs10<-rnorm(n=30, mean=40, sd=0.7)
obs11<-rnorm(n=30, mean=40, sd=0.7)
obs12<-rnorm(n=30, mean=40, sd=0.7)

observers1<-data.frame(cbind(obs1,obs2,obs3, obs4, obs5, obs6, obs7, obs8, obs9, obs10, obs11, obs12))

observers2<-data.frame(cbind(obs1,obs2,obs3, obs4, obs5, obs6, obs7, obs8, obs9, obs10, obs11, obs12))

observers<-list(observers1, observers2)
names(observers) <-c("obs1", "obs2")
```
###primary model
```{r}
umf1 <- unmarkedMultFrame(y = sampling.ocup1, 
                            siteCovs = data.frame(sampling.cov), 
                            yearlySiteCovs=data.frame(temps),
                            obsCovs=observers, numPrimary=3)
```
##Dynamic model
first term static variables
second term colonization (variable)
third extintion (variable)
detection (observer)
###the best model should take into acount v.1 and v.2, but not v.3, it shouldn't take into acount observers

```{r}
model1 <- colext(~v.1+v.2+v.3, ~1, ~1, ~1, umf1)

model2 <- colext(~1, ~1, ~1, ~1, umf1)

model3 <- colext(~v.1+v.2, ~1, ~1, ~1, umf1) #this should be the best model
model1
model2
model3

```
###graph
